---
title: "Meta-analysis psychoterapies and depression"
author: "Irene Alfarone"
date: "2025-05-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Dataset
Data are gathered from Cujipers et al. (2010) published meta-analysis on the differences between CR and SR depression measurements. The authors collect data from 48 studies that measure depression on both a clinician rating (HRSD) and self-report scale (BDI). The meta-analysis highlights a substantial difference between the patients’ and clinicians’ evaluations of depression, in favor of the clinician rating.

## Load data and clean dataset
```{r data}
library(readxl)
library(dplyr)
library(mixmeta)
library(tmvtnorm)
library(ggplot2)

d = read_xlsx("cujipersetal-data2.xlsx", sheet = "Sheet2")
head(d)

d <- d %>%
  mutate(
    across(c(-Study, -Psychotherapy), as.numeric),
    MissHRSD=ifelse(is.na(`T-HRSD-M-Post`), "0", "1"),
    MissBDI=ifelse(is.na(`T-BDI-M-Post`), "0", "1"),
    across(c(Psychotherapy), as.factor)
  )
head(d)

df <- d[-c(2, 4, 7, 13, 17, 19, 21, 27, 31, 34, 36, 37, 38, 45, 48, 56, 58, 59),]
head(df)

df$EstSR <- df$`T-BDI-M-Post` - df$`C-BDI-M-Post`
df$EstCR <- df$`T-HRSD-M-Post` - df$`C-HRSD-M-Post`

Spooled <- function(Nt, sdt, Nc, sdc)
{
  sqrt(((Nt-1)*sdt^2 + (Nc-1)*sdc^2)/((Nt-1)+(Nc-1)))
}

BDI_sp <- Spooled(df$Nt, df$`T-BDI-SD-Post`, df$Nc, df$`C-BDI-SD-Post`)
df$SESR <- BDI_sp * sqrt((1/df$Nt)+(1/df$Nc))


HRSD_sp <- Spooled(df$Nt, df$`T-HRSD-SD-Post`, df$Nc, df$`C-HRSD-SD-Post`)
df$SECR <- HRSD_sp * sqrt((1/df$Nt)+(1/df$Nc))

head(df)

dmnar = df[,c("Study", "N", "EstCR", "SECR", "EstSR", "SESR")]
dmnar$Cor.ws = 0.7
dmnar$Cor.ws[is.na(dmnar$EstCR) | is.na(dmnar$EstSR)] = NA
head(dmnar)
```

## Function for imputing data
```{r functions}
genimp.mnar = function(df, distribution = c("uniform", "normal", "tmvn"),
                       iter, minCR, maxCR, minSR, maxSR, meanCR, meanSR, sdCR, sdSR,
                       meantmv, sigmatmv, lower, upper, imprho, scaleSE) {
  distribution = match.arg(distribution)
  results = vector(mode = "list", length = iter)
  
  all_impCR = list()
  all_impSR = list()
  
  cor2cov = function(sd1, sd2, rho) {
    sd1 * sd2 * rho
  }
  
  for (i in 1:iter) {
    dfi = df 
    
    NmissCR = sum(is.na(dfi$EstCR))
    NmissSR = sum(is.na(dfi$EstSR))
    
    if (distribution == "uniform") {
      impCR = runif(NmissCR, min = minCR, max = maxCR)
      impSR = runif(NmissSR, min = minSR, max = maxSR)
      
    } else if (distribution == "normal") {
      impCR = rnorm(NmissCR, mean = meanCR, sd = sdCR)
      impSR = rnorm(NmissSR, mean = meanSR, sd = sdSR)
      
    } else if (distribution == "tmvn") {
      imputed = rtmvnorm(
        n = sum(is.na(dfi$Cor.ws)),
        mean = meantmv,
        sigma = sigmatmv,
        lower = lower,
        upper = upper
      )
      impCR = imputed[, 1]
      impSR = imputed[, 2]
    }
    
    all_impCR[[i]] = impCR
    all_impSR[[i]] = impSR
    
    dfi$EstCR[is.na(dfi$EstCR)] = sample(impCR, NmissCR, replace = FALSE)
    dfi$EstSR[is.na(dfi$EstSR)] = sample(impSR, NmissSR, replace = FALSE)
    dfi$Cor.ws[is.na(dfi$Cor.ws)] = imprho
    
    impSECR = sample(dfi$SECR[!is.na(dfi$SECR)], NmissCR, replace = TRUE) * scaleSE
    impSESR = sample(dfi$SESR[!is.na(dfi$SESR)], NmissSR, replace = TRUE) * scaleSE
    dfi$SECR[is.na(dfi$SECR)] = impSECR 
    dfi$SESR[is.na(dfi$SESR)] = impSESR 
    
    theta = cbind(dfi$EstCR, dfi$EstSR)
    Sigma = cbind(dfi$SECR^2,
                  cor2cov(dfi$SECR, dfi$SESR, dfi$Cor.ws),
                  dfi$SESR^2)
    
    mv = mixmeta(theta, Sigma, method = "reml")
    ci = confint(mv)
    
    results[[i]] = data.frame(
      iter = i,
      eff1 = mv$coefficients[1],
      eff2 = mv$coefficients[2],
      se1 = sqrt(mv$vcov[1, 1]),
      se2 = sqrt(mv$vcov[2, 2]),
      cov = mv$vcov[1, 2],
      ci.lb1 = ci[1, 1],
      ci.ub1 = ci[1, 2],
      ci.lb2 = ci[2, 1],
      ci.ub2 = ci[2, 2]
    )
  }
  
  return(list(
    impCR = unlist(all_impCR),
    impSR = unlist(all_impSR),
    results = do.call(rbind, results)
  ))
}

```

## Impute data
```{r imp}
###################
##### Uniform #####
###################

resuni = genimp.mnar(
  df = dmnar,
  distribution = "uniform",
  iter = 1000,
  minCR = -30,
  maxCR = 30,
  minSR = -30,
  maxSR = 30,
  imprho = 0.7,
  scaleSE = 1.5
)
resuni = resuni$results

##################
##### Normal #####
##################

resnorm = genimp.mnar(
  df = dmnar,
  distribution = "normal",
  iter = 1000,
  meanCR = 0,
  meanSR = 0,
  sdCR = 6,
  sdSR = 6,
  imprho = 0.7,
  scaleSE = 1.5
)

hist(resnorm$impCR, col = "#C64191", freq = FALSE,
     main = "All Imputed EstCR",
     xlab = "EstCR")
curve(dnorm(x, mean=0,sd=6), add=T, col="#028090")

hist(resnorm$impSR, col = "#F0F3BD", freq = FALSE,
     main = "All Imputed EstSR",
     xlab = "EstSR")
curve(dnorm(x, mean=0,sd=6), add=T, col="#028090")

resnorm = resnorm$results

resnorm2 = genimp.mnar(
  df = dmnar,
  distribution = "normal",
  iter = 1000,
  meanCR = 6,
  meanSR = 6,
  sdCR = 6,
  sdSR = 6,
  imprho = 0.7,
  scaleSE = 1.5
)

hist(resnorm2$impCR, col = "#C64191", freq = FALSE,
     main = "All Imputed EstCR",
     xlab = "EstCR")
curve(dnorm(x, mean=6,sd=6), add=T, col="#028090")

hist(resnorm2$impSR, col = "#F0F3BD", freq = FALSE,
     main = "All Imputed EstSR",
     xlab = "EstSR")
curve(dnorm(x, mean=6,sd=6), add=T, col="#028090")

resnorm2=resnorm2$results

resnorm3 = genimp.mnar(
  df = dmnar,
  distribution = "normal",
  iter = 1000,
  meanCR = -6,
  meanSR = -6,
  sdCR = 10,
  sdSR = 12,
  imprho = 0.7,
  scaleSE = 1.5
)

hist(resnorm3$impCR, col = "#C64191", freq = FALSE,
     main = "All Imputed EstCR",
     xlab = "EstCR")
curve(dnorm(x, mean=-6,sd=10), add=T, col="#028090")

hist(resnorm3$impSR, col = "#F0F3BD", freq = FALSE,
     main = "All Imputed EstSR",
     xlab = "EstSR")
curve(dnorm(x, mean=-6,sd=12), add=T, col="#028090")

resnorm3 = resnorm3$results

#############################
##### Truncate MVNormal #####
#############################

lower = c(-Inf, -Inf) # for now simply multivariate
upper = c(Inf, Inf)

restmvn = genimp.mnar(
  df = dmnar,
  distribution = "tmvn",
  iter = 1000,
  lower = lower,
  upper = upper, 
  meantmv = c(-6,-6),
  sdCR = 6,
  sdSR = 6,
  sigmatmv = matrix(c(6^2, 0.7 * 6 * 6, 0.7 * 6 * 6, 6^2), nrow = 2), 
  imprho = 0.7,
  scaleSE = 1.5
)

restmvn = restmvn$results
```

## Function for summarizing results
```{r sum}
sum.meth = function(res, true1, true2, method_name) {
  m = nrow(res)
  
  Q_mat = cbind(res$eff1, res$eff2)
  Q_bar = colMeans(Q_mat)                     
  
  B = cov(Q_mat)                              
  
  U_list = lapply(1:m, function(i) {
    matrix(c(res$se1[i]^2, res$cov[i], 
             res$cov[i], res$se2[i]^2), 
           nrow = 2)
  })
  
  U_bar = Reduce("+", U_list) / m             
  
  Tmat = U_bar + (1 + 1/m) * B # total pooled variance (now I incorporate also the covariance)
  
  se = sqrt(diag(Tmat))
  
  df = (m - 1) * (1 + diag(U_bar) / ((1 + 1/m) * diag(B)))^2
  
  ci1 = Q_bar[1] + c(-1, 1) * qt(0.975, df[1]) * se[1]
  ci2 = Q_bar[2] + c(-1, 1) * qt(0.975, df[2]) * se[2]
  
  
  return(data.frame(
    method = method_name,
    est_CR = Q_bar[1],
    est_SR = Q_bar[2],
    pci_lb_CR = ci1[1],
    pci_ub_CR = ci1[2],
    pci_lb_SR = ci2[1],
    pci_ub_SR = ci2[2]
  ))
}

results = rbind(
  sum.meth(resuni, true1, true2, "Uniform"),
  sum.meth(resnorm, true1, true2, "Normal(0,0)"),
  sum.meth(resnorm2, true1, true2, "Normal(6,6)"),
  sum.meth(resnorm3, true1, true2, "Normal(-6,-6)"),
  sum.meth(restmvn, true1, true2, "TMVN")
)

```


## Inspect results
```{r insp}
print(results)

resuni$method = "Uniform"
resnorm$method = "Normal(0,0)"
resnorm2$method = "Normal(6,6)"
resnorm3$method = "Normal(-6,-6)"
restmvn$method = "TMVN"

results.table = rbind(resuni, resnorm, resnorm2, resnorm3, restmvn)
head(results.table)
```

## Plot results
```{r plot}
ggplot(results, aes(x = method)) +
  geom_linerange(aes(ymin = pci_lb_CR, ymax = pci_ub_CR), linewidth = 1.5) +
  geom_point(aes(y = est_CR), color = "black", size = 3) +
  geom_hline(yintercept = -6.0604, linetype = "dashed", color = "#A4031F") +
  coord_flip() +
  labs(
    title = "Uncertainty Intervals for CR",
    x = "Method", y = "CR Interval"
  ) +
  theme_minimal()


ggplot(results, aes(x = method)) +
  geom_linerange(aes(ymin = pci_lb_SR, ymax = pci_ub_SR), linewidth = 1.5) +
  geom_point(aes(y = est_SR), color = "black", size = 3) +
  geom_hline(yintercept = -6.4751, linetype = "dashed", color = "#A4031F") +
  coord_flip() +
  labs(
    title = "Uncertainty Intervals for SR",
    x = "Method", y = "CR Interval"
  ) +
  theme_minimal()
```
